---
title: "Exam 1 Template"
author: "YOUR NAME HERE"
date: "DATA 2401 - Fall 2022"
output:
  html_document: default
  pdf_document: default
---



> “We will be honest in all our academic activities and will not tolerate dishonesty.” 

> – UHD Academic Honor Code


&nbsp;

&nbsp;

You are submitting this exam under the UHD honor code. Answer all below questions. Complete the questions, knit the file to html, commit and push to your repository *BEFORE 2:45* and turn in link to the repository to Blackboard. __Any commit made after 2:45 pm will not be graded.__ 

\newpage 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




### Part 1 (20 points)

* You will get 10 points of credit for a properly knit RMD/html(or pdf) file.
* You will get 10 points of credit for a properly committed and pushed GitHub repository, with clear and descriptive commit descriptions. 

## Part 2 (10 points)

In your own words, *briefly* describe the difference between saving, committing, and pushing. 
saving is saving the work on r studio . committing is adding comments and notes of changes made to the r project. Push is pushing the work back on to the repository

## Part 3 (10 points)

* [2 points] Create a vector named `numbers` that contains the odd numbers from 3 to 15. 
```{r}
numbers <-seq(3,15, by = 2)
numbers
```


* [3 points] Create a second vector called `half_numbers` that contains the elements of `numbers` divided by 2
```{r}
half_numbers <-numbers /2
half_numbers
```

* [2 points] Add the string "puppies" to the end of `numbers`. 
```{r}
numbers <- seq(3,15, by = 2)
numbers <- c(numbers, "puppies")
```

* [3 points] Try to divide the new vector by two. Does it work? Why or why not? 

```{r}
numbers <- seq(3,15, by = 2)
numbers_list <- c(numbers, "puppies")
half_numbers <-as.numeric(numbers_list[-length(numbers_list)])/2
half_numbers
```

## Part 4 (15 points)

You run a small dog daycare. This is the data you have on the dogs that come to your business.

```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
weight <- c(14, 89, 32, 46, 55, 7, 25, 36)
```

* [2 points] Find the average weight and age of the dogs.
```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
weight <- c(14, 89, 32, 46, 55, 7, 25, 36)
average_age <- mean(age)
average_weight <- mean(weight)
average_age
average_weight
```

* [2 points] How light is the dog with the lowest weight?
```{r}
 min(weight)
```

* [3 points] What is the name of the oldest dog?
```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
weight <- c(14, 89, 32, 46, 55, 7, 25, 36)
names[which.max(age)]
age[which.max(age)]
```

* [3 points] What is the age of the heaviest dog?
```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
weight <- c(14, 89, 32, 46, 55, 7, 25, 36)
age[which.max(weight)]

```

* [2 points] It's been 2 years since you entered this information! Change `age` to
reflect that. 
```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
age2 <- age +2
age2

```

* [3 points] A new dog is here! Her name is Daisy, she is 3 years old, and weighs 39 pounds. Add her to your data. 
```{r}
names <- c("Fido", "Doug", "Bella", "Max", "Cooper", "Lucy", "Duke", "Barkley")
age <- c(2, 5, 4, 6, 1, 13, 9, 3)
weight <- c(14, 89, 32, 46, 55, 7, 25, 36)
names <- c(names, "daisy")
age <- c(age, 3)
weight <- c(weight, 39)
data.frame(names, age, weight)
```


## Part 5 (10 points)

In baseball, the slugging percentage is computed with 5 variables, the number of singles, doubles, triples, homeruns and at-bats. We abbreviate these 1B, 2B, 3B, HR and AB. 

1. [7 points] Write a function that takes these 5 values as arguments and returns the slugging percentage for that player. $$  \frac{(1B + 2B\times2 + 3B \times 3 + HR\times 4)}{AB} $$
```{r}
slugging_percentage <- function(singles, doubles, triples, home_runs, at_bats) {
    if(at_bats == 0){
      return(0)
    }
    total_bases <-singles + (doubles *2) +
      (triples * 3) + (home_runs * 4)
    slg <- total_bases / at_bats
    return(slg)
}
slugging_percentage(50, 20, 5, 10 ,200)
```



2. [3 points] The Astros player Jose Altuve currently has the following stats: 

`1B = 89, 2B = 34, 3B = 0, HR = 25, AB = 499 `

Use your function to find his slugging percentage. 
```{r}
slugging_percentage <- function(singles, doubles, triples, home_runs, at_bats) {
    if(at_bats == 0){
      return(0)
    }
    total_bases <-singles + (doubles *2) +
      (triples * 3) + (home_runs * 4)
    slg <- total_bases / at_bats
    return(slg)
}
slugging_percentage(89, 34, 0, 25, 499)
```


## Part 6 (15 points)

1. [6 points] Write a function that takes a vector `x` and returns a vector with the square root of each element of the vector. 

> For example, if you pass the function a vector containing 4 and 16, it should return a vector containing 2 and 4. 

```{r}
sqrt_vector <- function(x){
  return(sqrt(x))
}
my_vector <- c(4, 9, 16, 25)
sqrt_vector(my_vector)
```


2. [6 points] Write a second function that does the above, but also removes any negative values or changes them to 0.
```{r}
sqrt_vector_no_neg <- function(x){
  x[x < 0] <-0
  return(sqrt(x))
}
my_vector <- c(4, -9, 16, -25, 0)
sqrt_vector_no_neg(my_vector)
```


> For example, if you pass *this* function -4 and 16, it should return either 0 and 4 or just 4.

3. [3 points] Test your second function on the following vectors:

```{r}
test1 <- c(1:10)
test2 <- c(-4:4)
# Function: square root, negatives replaced with 0
sqrt_vector_no_neg <- function(x) {
  x[x < 0] <- 0
  return(sqrt(x))
}

# Test vectors
test1 <- c(1:10)
test2 <- c(-4:4)

# Apply function and print results
cat("Square roots with negatives as 0 for test1:\n")
print(sqrt_vector_no_neg(test1))

cat("\nSquare roots with negatives as 0 for test2:\n")
print(sqrt_vector_no_neg(test2))
```


## Part 7 (10+3 points)

A fancy tech company gives employees ratings of 1, 2 and 3, which are used to compute their annual bonuses. If an employee gets a score of 3, their bonus is 15% of their salary, if they get a 2, their bonus is 10% of their salary, if they get a 1, their bonus is 5% of their salary. 

Rating | Bonus Percent
:-----:|:-----:
3|15%
2|10%
1|5%

1. [8 points] Write a function that takes the employees score and their salary as arguments and returns their bonus amount for the year. 
```{r}
# Function to calculate annual bonus
annual_bonus <- function(score, salary) {
  if(score == 3) {
    bonus <- 0.15 * salary
  } else if(score == 2) {
    bonus <- 0.10 * salary
  } else if(score == 1) {
    bonus <- 0.05 * salary
  } else {
    bonus <- 0 # In case of invalid score
  }
  return(bonus)
}
# Example usage:
annual_bonus(3, 60000) # Returns 9000
annual_bonus(2, 60000) # Returns 6000
annual_bonus(1, 60000) # Returns 3000
```

2. [2 points] Fitz makes $100,000 a year, but only got a score of 1. Use your function to calculate his bonus.
```{r}
annual_bonus <- function(score, salary) {
  if(score == 3) {
    bonus <- 0.15 * salary
  } else if(score == 2) {
    bonus <- 0.10 * salary
  } else if(score == 1) {
    bonus <- 0.05 * salary
  } else {
    bonus <- 0 # In case of invalid score
  }
  return(bonus)
}
annual_bonus(1,100000)

```





3. [3 extra credit points] Modify your function so that it returns the full pay for the year, salary and bonus combined. How much will Fitz make in total?
```{r}
# Function to calculate total pay including bonus
total_pay <- function(score, salary) {
  if(score == 3) {
    bonus <- 0.15 * salary
  } else if(score == 2) {
    bonus <- 0.10 * salary
  } else if(score == 1) {
    bonus <- 0.05 * salary
  } else {
    bonus <- 0 # For invalid score
  }
  
  total <- salary + bonus
  return(total)
}

# Example usage for Fitz:
total_pay(1, 100000)

```
